---
title: "Water Quality Labels and Datasheets"
author: "Emily Nussdorfer and Katie Wampler"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("helper-funs.R")
require("openxlsx2")
require("here")
```

# 1. Input Sampling Info

## Study and Site Info

-   **`study`:** 4-5 letter code associated with the study
-   **`samp_type:`** type of sample being collected
    -   `GB`: Grab samples
    -   `IS`: ISCO samples
    -   `SS`: Soil sample
-   **`rn`:** The number to start the random number sequence. This is used to assign a random analysis order.

```{r study-info}
study <- "BDRK"
samp_type <- "IS"
rn <- 1
```

## Sampling Info

### Grab Samples

*(only required if `samp_type` is "GB")*

-   **`sites`:** One of the following options:

    -   File path to a `.csv` of site names where the first column is the site names.

    -   To leave the site names blank, use `"______"` instead.

-   **`start`:** Either a line for a fully blank label or a date followed by a line for the time (ie. `"YYYYMMDD____"`)

```{r grab-sample-info}
if(samp_type == "GB"){
  sites <- "example_sites.csv"
  start <- "____________" 
}
```

### ISCO Samples

*(only required if `samp_type` is "IS")*

-   **`sites`:** A vector of site names (ie. `c("STA","COA")`)

-   **`nsamp`**: The number of samples that will be collected at each site. Should either be a single number if consistent across sites, or formatted as a vector the same length as the number of sites (ie. `c(24,24)`)

-   **`interval`**: The number of hours between each sample taken. Should either be a single number if consistent across sites, or formatted as a vector the same length as the number of sites (ie. `c(2,2)`)

-   **`start`:** The start time to use on the labels, either:

    -   A vector with the start time(s) if known formatted as YYYY-MM-DD HH:MM. Should either be a single character if consistent across sites, or formatted as a vector the same length as the number of sites (ie. `c("2026-02-23 15:00","2026-02-23 17:00")`)

    -   If the start time is unknown use either a line for a fully blank label or a date followed by a line for the time (ie. `"YYYYMMDD____"`)

```{r isco-sample-info}
if(samp_type == "IS"){
  sites <- c("STA","COA")
  nsamp <- c(24,24)
  interval <- c(2,2)
  start <- c("2026-02-23 15:00","2026-02-23 17:00")
  #start <- "20260223____"
}
```

## Blanks

**There are two types of blanks you may want to include:**

-   **laboratory blanks (`nLB`)**: Blanks that test for contamination throughout the sample processing steps.

-   **field blanks (`nFB`)**: Blanks that test for contamination throughout the entire sample collection process in the field through the analysis.

Specify the number to include below.

```{r blanks}
nLB <- 2
nFB <- 0
```

## Analysis Info

Logical (`TRUE`/`FALSE`) for which analysis will be performed on the samples:

-   **`DOM`:** dissolved organic matter composition on the Aqualog
-   **`DOC`:** dissolved organic carbon on the Shimadzu
-   **`CCAL`:** additional analytes sent to outside labs
-   **`Levoglucosan`:** levoglucosan to be archived for future analysis
-   **`BPCA`:** benzenepolycarboxylic acid to quantify black carbon
-   **`Excess`:** additional sample collected for other purposes (ie., sending to Node 3)

```{r analysis-info}
DOM <- TRUE
DOC <- TRUE
CCAL <- FALSE
Levoglucosan <- FALSE
BPCA <- FALSE
Excess <- FALSE
```

## Export Location

Specify where to export the spreadsheet with the sample names and datasheets.

**`saveloc`:** the filepath to the location to save the excel file (`.xlsx`) to.

```{r export-info}
saveloc <- "example-sample-labels.xlsx"
```

# 2. Create Labels

## Create the SampleIDs

Before creating the labels we will create the sample ID's based on the provided information.

!!!! bottle number for isco??

```{r generate-ids include=FALSE}
set.seed(9) #keep the randomization the same each time the code is run. 

generate_ids(study, samp_type, sites, start, nLB, nFB, rn, nsamp, interval)
# !!! write to excel template (Sample-DateTime) 
  #!!! only overwrite if blank, if not, just create the new filled IDs
```

```{r create-labels include=FALSE}
# !!!! check for if datetimes are in spreadsheet, if yes run otherwise don't 
# section 6-8 here
# !!! write to excel template (Labels-Field & Labels-Lab)
```

```{r create-datasheets include=FALSE}
# !!!! check for if datetimes are in spreadsheet, if yes run otherwise don't 
# !!! write to excel template (DS-Filter Weight, DS-Bottle Numbers, DS-ISCO Bottle Weight)
```
