---
title: "wq-sample-label-datasheet-templates"
output:
  pdf_document: default
  header-includes:
    - \usepackage{pdflscape}
    - \newcommand{\blandscape}{\begin{landscape}}
    - \newcommand{\elandscape}{\end{landscape}}
---

# Summary

The purpose of this R markdown is to ...

Some packages that might be of use:

```{r}
library(dplyr)
library(stringr)
library(kableExtra)
```

The dplyr package is specifically designed to do data manipulation in a way that's relatively easy to read using things called pipes (`%>%`). See the chapter in this textbook [here](https://r4ds.hadley.nz/data-transform.html) for some examples of using `dplyr`.

# 1. User Inputs

**Put info from excel FILLME here as variables. Since we updated the methods we'll probably need to add:**

-   `random_start` for the starting random number

-   `samp_type` (GB,IS, SS, etc.)

Variable names should be lowercase with no spaces (use underscores instead) and not too long (abbreviate where it makes sense).

```{r}
project <- "WWS" #the large project code
study <- "TEST" #the study code 
```

# 2. Bottle Calculator

Based on number of sites and samples per site, calculate bottle types needed as data.frame, then we can print nicely.

```{r}

```

# 3. Label IDs

We'll use this to generate sample IDs and random IDs for labels. For a start just use R functions to take the info from 1 and convert to the IDs. I think you have the most updated things we need but I think we need:

-   **SampleID:** project_site_type_date_time

-   **RandomID:** project_r###

-   **ISCOID:** just the number order it goes in the isco?

```{r}
#paste function works well to join things together 
paste0("A", "B", "C")

paste("A", "B", "C")

paste("A", "B", "C", sep="_")

#the seq function is good for generating a sequence 
seq(1,4)

seq(as.Date("2023-02-01"), as.Date("2024-01-03"), by="month")

#we can store all this in a data.frame similar to excel 
data <- data.frame(project = project, 
                   study=study, 
                   dates = seq(as.Date("2023-02-01"), 
                               as.Date("2024-01-03"), by="month"))

data
```

# 4. Specifying analyte

We might need to do some experimenting here. For a set of ISCO samples or a small number of grab samples I think we could just let the user specify the analytes they plan to run (i.e., `analyte <- c("DOC", "DOM")`). If we want to get fancier and let people run only a subset of samples for something we'll need to get fancier. But for now maybe we just allow all or nothing.

# 5. Generating the .csv for labels

We'll want to write the labels we need to a .csv file so we can use it for making labels. We'll want to just pull the columns we want.

```{r}
print_data <- data %>% select(project, study)
print_data
#write.csv(print_data, "filename.csv", row.names=FALSE)
```

# 6. Making the tables

This part will be a little more complicated, since we want specific formatting. I think if we use the LaTeX command below (which is what the Rmarkdown uses to convert to PDF), it should start a new page that we can format how we like. I'm new to this part too, so we could do some research together when you get here.

\\newpage

\blandscape

Date: \_\_\_\_\_\_\_\_\_\_\_\_\_

Technician: \_\_\_\_\_\_\_\_\_\_\_\_\_\_

Project Code: \_\_\_\_\_\_\_\_\_\_\_\_\_\_

Study Code: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

```{r echo=FALSE}
kable(data, booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position"), 
                full_width = FALSE, 
                font_size = 10) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#D7261E")
```

\elandscape
