---
title: "wq-sample-label-datasheet-templates"
output:
  pdf_document: default
  header-includes:
    - \usepackage{pdflscape}
    - \newcommand{\blandscape}{\begin{landscape}}
    - \newcommand{\elandscape}{\end{landscape}}
---

```{r libraries}
library(dplyr)
library(stringr)
library(kableExtra)
```

# Importing data

```{r}
samplesheet <- read.csv("example_label-table2.csv")
```

# Making the tables

This part will be a little more complicated, since we want specific formatting. I think if we use the LaTeX command below (which is what the Rmarkdown uses to convert to PDF), it should start a new page that we can format how we like. I'm new to this part too, so we could do some research together when you get here.

\\newpage

\blandscape

Date: \_\_\_\_\_\_\_\_\_\_\_\_\_

Technician(s): \_\_\_\_\_\_\_\_\_\_\_\_\_\_

```{r bottle numbers, echo=FALSE}
kable(samplesheet,booktabs=TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position","repeat_header"), 
                full_width = FALSE, 
                font_size = 10) %>%
  row_spec(0, bold = TRUE, color = "black", background = "#67c4db")

```

\elandscape

\newpage

```{r bottle weights, echo=FALSE}
kable(samplesheet,booktabs=TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position","repeat_header"), 
                full_width = FALSE, 
                font_size = 10) %>%
  row_spec(0, bold = TRUE, color = "black", background = "#67c4db")
```

\newpage

```{r filter weights, echo=FALSE}
kable(samplesheet,booktabs=TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position","repeat_header"), 
                full_width = FALSE, 
                font_size = 10) %>%
  row_spec(0, bold = TRUE, color = "black", background = "#67c4db")

```
