---
title: "Untitled"
author: "Emily Nussdorfer"
date: "2026-07-01"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

Filling in pt 1 #this is for unique site names! #this one is assuming
you are going to the field with blank times and will fill the dates into
the generated spreadsheet use that to print lab labels

## **Section 1- include sample information ([editable]{.underline}):**

```{r}
######### Filling in codes##########
study <- "BDRK" #the study code
samp_type <- "GB" #(GB,IS, SS, etc.)
rn <- 1 #the start of a random number sequence. May change with projects or studies
site <- read.csv("example_sites.csv") #import your site list. This file needs to have the header "site"
#not sure how to do this without a specific header...
site
blank_time <- "____________" #either have it blank or with sampling date "____________" or "YYYYMMDD____"
blank_time
sc <- "20260823" #what sample campaign this is. Either date ex."20260223 or campaign ex. "1"

#fill out analysis
DOM <- TRUE
DOC <- TRUE
CCAL <- FALSE
Levoglucosan <- TRUE
Excess <- FALSE

####optional formatting- fill out#####
alldata_blank_file <- "alldata_blank" #what to call header for blank times alldata file
alldata_file <- "alldata" #what to call header for alldata file you fill out
field_file <-  "fieldsamples" #what to call header for field label file
lab_file <-  "labsamples" #what to call header for lab label file

#what format the files will end with. ex this returns:
fileformat <- paste0(study,"_",samp_type,"_",sc,"_label-table.xlsx") 
blank_fileformat <- paste0("blank_",study,"_",samp_type,"_",sc,"_label-table.csv") 

```

## **Section 2- run libraries:**

```{r}
library(dplyr)
library(stringr)
library(lubridate)
library(writexl)
```

## **Section 3- code:**

Includes: total sample amount, and sample info, and functions to prevent
overwrite files

```{r}
sn <- nrow(site) #how many sites there are

#safewrite function 
safe.write_xlsx <- function(x, path, ...) {
  if(file.exists(path)) {
    stop("The file you are trying to write already exists. DO NOT OVERWRITE UNLESS YOU MEAN TO!\n",
         "We don't want to overwrite randomized sample names\n",
         "Check ", path, " to see if the already written file is ok.")
  }
  write_xlsx(x, path, ...)
}
safe.write_csv <- function(x, path, ...) {
  if(file.exists(path)) {
    stop("The file you are trying to write already exists. DO NOT OVERWRITE UNLESS YOU MEAN TO!\n",
         "We don't want to overwrite randomized sample names\n",
         "Check ", path, " to see if the already written file is ok.")
  }
  write.csv(x, path, ...)
} 
```

## **Section 4- code:**

Saves a file that will be used to print off field labels. It will need
to be edited later before printing lab data sheets and labels

-   Remember to save your code

-   Before proceeding to section 5, you will need to fill out the
    created spreadsheet columns with the date and times

```{r warning=TRUE}
#Adds random numbers to data frame. Re-running will randomize them again!
datablank<- data.frame(site) %>%
  mutate(study=study,
         samp_type=samp_type,
         blank_time=blank_time,
         randomn= sample(rn:sn,sn, replace=F),
         date_text_PST = "",
         time_text_PST ="") 
datablank

#adds blank sample IDs, sample names, and adds text columns to fill out for date and time
blank_labels <- datablank %>%
  mutate(sample_name=paste0( study, "_", "R", sprintf("%04d", randomn)), #adds sample name
         field_sample_ID_blank = paste0(study, "_",
                                        site, "_",
                                        samp_type,"_",
                                        blank_time))#adds blank sample ID
         #date_text_PST = rep('', nrow(datablank)), time_text_PST = rep('', nrow(datablank))) #adds column names to fill out could've been added in previous code section but I like it being on the far right for data entry. Do later?
blank_labels 
#print blank labels 
print <- blank_labels %>% select(
  sample_name,field_sample_ID_blank,randomn,study,samp_type,site,date_text_PST,time_text_PST) #reorder so it's easier for data entry
safe.write_csv(print, blank_fileformat, row.names=FALSE)

stop("SAVE YOUR CODE.Before printing lab labels out- fill out blank columns for date and time and print lab labels (as text only! example: for 2PST write 1400- make sure it's not PDT), save file.")


```

## **Section 5 - code**

Before running this part:

-   fill out the time_text column and save file

-   rerun sections 1-3. Do [not]{.underline} rerun section 4

```{r edit, warning=TRUE}
datalabels <- read.csv(blank_fileformat) #bring in the saved data labels. 
datalabels
stop("check that date_text and time_text rows were filled out before proceeding")
```

## **Section 6:**

```{r}
#make datetimes and the sampleID with times

data<-datalabels %>%
  mutate(datetime_PST=
           paste0(sprintf("%06d",datalabels$date_text_PST),"_",
                  sprintf("%04d",datalabels$time_text_PST)),
         field_sample_ID =
           paste0(study, "_",
                  site, "_",
                  samp_type,"_",
                  datetime_PST))

data
#sort the data by the randomn, since they should be filtered that way
datasortlab <- data[order(data$randomn),]
datasortlab #use this to print datasheets out
```

## **Section 7:**

```{r}
#add analytes that are TRUE
analysis_opt <- c("CN", "AQ", "NU", "EX", "LV")
analysis_choice <- c(DOC, DOM, CCAL, Excess, Levoglucosan)

analysiscodes <- analysis_opt[analysis_choice]
analysiscodes

analysisn <- length(analysiscodes) #how many analytes there are
samples <- analysisn*sn #how many total filtered samples

datacodes<- data.frame(code = rep(analysiscodes,times=sn)) %>% #add labcodes
  mutate(analysis = case_when(
    code == "CN" ~ "Shimadzu",
    code == "AQ" ~ "Aqualog",
    code == "NU" ~ "CCAL",
    code == "LV" ~ "Levoglucosan",
    code == "EX" ~ "Excess",
    TRUE ~ NA_character_
  ))
datacodes
#repeat samples for each analysis
labdata <- datasortlab %>% slice(rep(1:n(), each = analysisn)) %>%
mutate(datacodes,
       lab_sample_ID =
        paste0(study, "_",
                      site, "_",
                      samp_type,"_",
                      datetime_PST,"_", 
                      datacodes$code ))

labdata
```

## **Section 8- printing:**

```{r}
write_data.frames <- list(
 
  lab_labels   = labdata %>% select(sample_name,lab_sample_ID,analysis),
  lab_filter_sheets = datasortlab %>% select(sample_name,field_sample_ID),
  label_metadata = data %>% select(site,datetime_PST,study,samp_type,randomn) #was just a dump for things
)
safe.write_xlsx(write_data.frames,
                path = fileformat)

```
