---
title: "ISCO code"
author: "Emily Nussdorfer"
date: "2026-07-01"
---

Known time samples: Run code as usual

Unknown time samples: Sections 1-4

note: could combine sections 1-3 into one part

## **Section 1- include sample information ([editable]{.underline}):**

```{r sample info, include=FALSE}

#sample information:
study <- "SSEMG" #the study code
samp_type <- "IS" #"IS" = ISCO
rn <- 1 #the start of a random number sequence. May change with projects or studies
site <- c("STA","COA") #sites as a vector
nsamp <- c(24,24) #amount of samples per site. Matches the site vector above.
interval <- c(2,2) #interval of sampling ISCO is programmed for, usually 2hrs
sc <- "20260223" #what sample campaign this is, the date it starts "YYYYMMDD" or something like sample_campaign-#. This is for file name purposes and doesn't effect sample naming

#sample times:
knowntime <- TRUE #fill TRUE When you have a time start and FALSE for if you don't know the time start. Knowing the time is the most efficient method.Important for If statements.  
#fill if you have known time:
start <- c("2026-02-23 15:00","2026-02-23 17:00")
#fill if you don't know time:
blanktime <- "____________" #either have it blank  "____________" or with sampling date "YYYYMMDD____" or "YYYYMM______", etc

#analysis, fill TRUE if using them, FALSE if not:
DOM <- TRUE
DOC <- TRUE
CCAL <- TRUE
Levoglucosan <- TRUE
Excess <- TRUE

#optional change:
#can add a folder here if desired
#what format the files will end with. ex this returns: study_sapmpletype_date_label-table 
fileformat <- paste0(study,"_",samp_type,"_",sc,"_label-table.xlsx") #the file name
fileformat
blank_fileformat <- paste0("blank_",study,"_",samp_type,"_",sc,"_label-table.csv") 
blank_fileformat 

```

## **Section 2- run libraries:**

```{r libraries, include=FALSE}
library(dplyr)
library(stringr)
library(lubridate)
library(writexl)
```

## **Section 3- code:**

Includes: total sample amount, and sample info, and functions to prevent overwrite files

```{r pt 1, include=FALSE}
sn <- sum(nsamp) #total sample amount from the vector that was filled out

#adding sample information
sampleinfo <- data.frame(site=site,nsamp=nsamp,start=start,interval=interval,study=study,samp_type=samp_type,blanktime=blanktime) #all sample info for the function
#would need to alter if it includes grabs 

# Functions to safely write slsxs and csvs without overwriting anything
safe.write_xlsx <- function(x, path, ...) {
  if(file.exists(path)) {
    stop("The file you are trying to write already exists. DO NOT OVERWRITE UNLESS YOU MEAN TO!\n",
         "We don't want to overwrite randomized sample names\n",
         "Check ", path, " to see if the already written file is ok.")
  }
  write_xlsx(x, path, ...)
}
safe.write_csv <- function(x, path, ...) {
  if(file.exists(path)) {
    stop("The file you are trying to write already exists. DO NOT OVERWRITE UNLESS YOU MEAN TO!\n",
         "We don't want to overwrite randomized sample names\n",
         "Check ", path, " to see if the already written file is ok.")
  }
  write.csv(x, path, ...)
} #had to add this because couldn't get blank files to work wth xlsx
```

## **Section 4- code:**

Creates data frame for the labels

-   Re-running the code will re-randomize the random numbers

-   Saves a csv file for blank time samples

```{r code pt 2, warning=TRUE, include=FALSE}
#function to fill out site, study, sample type, time stamps, and bottle numbers
#If statement, defines the site_label function depending on if there is a known time or not
if (knowntime == TRUE){
  #Known time 
  site_label <- function(sampleinfo){
    sitelist <- rep(sampleinfo$site, times = sampleinfo$nsamp) #if you turn into a df to return you wouldn't need this, it would automatically recycle
    timestart <- as.POSIXct(sampleinfo$start)
    timelist <- seq(timestart, by=paste0(sampleinfo$interval, " hours"), length = sampleinfo$nsamp)
    botnum <- rep(1:24, times= ceiling(sampleinfo$nsamp/24))[1:sampleinfo$nsamp] #added interval so if we ever want to change it, it's flexible
    return(data.frame(site =sitelist, timelist=timelist, 
                      botnum = sprintf("%02d",botnum), 
                      study=sampleinfo$study,
                      samp_type=sampleinfo$samp_type,
                      datetime_PST=format(timelist,"%Y%m%d%_%H%M"))) #could include other label things in here too
  }
} else if (knowntime == FALSE){
  site_label <- function(sampleinfo){
    sitelist <- rep(sampleinfo$site, times = sampleinfo$nsamp) #if you turn into a df to return you wouldn't need this, it would automatically recycle
    botnum <- rep(1:24, times= ceiling(sampleinfo$nsamp/24))[1:sampleinfo$nsamp] #added interval so if we ever want to change it, it's flexible
    return(data.frame(site =sitelist,
                      blank_time=sampleinfo$blanktime,
                      botnum = sprintf("%02d",botnum),
                      study=sampleinfo$study,samp_type=sampleinfo$samp_type)) #could include other label things in here too
  }
}







#make the sample table from the function and add random numbers to it. Re-running will change the random numbers.
data <- lapply(1:nrow(sampleinfo),function(x){site_label(sampleinfo[x,])})%>%
  bind_rows()%>%
  mutate (randomn= sample(rn:sn,sn, replace=F))

#add sample names and IDs
if (knowntime == TRUE){
  data_labels<-data %>%
    mutate(sample_name=paste0( study, "_", "R", sprintf("%04d", randomn)),
           field_sample_ID = paste0(study, "_",
                                    site, "_",
                                    samp_type,"_",
                                    datetime_PST))
}else if (knowntime == FALSE ){
  data_labels<-data %>%
    mutate(sample_name=paste0( study, "_", "R", sprintf("%04d", randomn)),
           field_sample_ID_blank = paste0(study, "_",
                                          site, "_",
                                          samp_type,"_",
                                          blanktime))}

#this prints your data labels out if you have blank samples, since you will need to rerun this code again. 
if(knowntime == FALSE){
  safe.write_csv(data_labels,
                 path = blank_fileformat,
                 row.names = FALSE)
  stop("DO NOT CONTINUE UNTIL YOU HAVE YOUR TIMES. SAVE YOUR CODE")}

```

## **Section 5 - code** (for [blank times only]{.underline}):

*Note: can skip to section 6 if you have known times*

Follow if you had blank times (this won't do anything for known times):

-   Do not change any of the above sections (leave known time FALSE)

-   Re-run sections 1-3 only

-   Fill the times and run code below ([editable]{.underline}):

```{r pt 5: and run, include=FALSE}
#enter the known start times on the ISCO (look at section 1 to match it to the correct site order)
newtimes <- c("2026-02-23 15:00","2026-02-23 17:00")
```

-   Run code below:

```{r pt 5: just run it, include=FALSE}
if(knowntime == FALSE){
  datablank <- read.csv(blank_fileformat)} #imports file using previous file naming

if(knowntime == FALSE){
  datablank <- read.csv(blank_fileformat) #imports file using previous file naming
  site_label_blank <- function(sampleinfo){
    timestart <- as.POSIXct(sampleinfo$start)
    timelist <- seq(timestart, by=paste0(sampleinfo$interval, " hours"), length = sampleinfo$nsamp)
    return(data.frame(time_PST=as.character(timelist),datetime_PST=format(timelist,"%Y%m%d%_%H%M"))) 
  } #make a data frame with times
  data_times <- lapply(1:nrow(sampleinfo),function(x){site_label_blank(sampleinfo[x,])}) %>% #makes the frame
  bind_rows%>%
  mutate(datablank) #add data from time-blank file
  data_labels <- mutate(data_times, field_sample_ID = paste0(study, "_",
                                                          site, "_",
                                                          samp_type,"_",
                                                          datetime_PST)) #add sample Ids with time 
  } 





```

## **Section 6- adding lab analytes:**

```{r include=FALSE}
#adding in processing information:
#add analytes that are TRUE
analysis_opt <- c("CN", "AQ", "NU", "EX", "LV")
analysis_choice <- c(DOC, DOM, CCAL, Excess, Levoglucosan)
analysiscodes <- analysis_opt[analysis_choice]
analysiscodes
analysisn <- length(analysiscodes) #how many analytes there are
samples <- analysisn*sn #how many total filtered samples there will be

datacodes<- data.frame(code = rep(analysiscodes,times=sn)) %>% #add labcodes
  mutate(analysis = case_when(
    code == "CN" ~ "Shimadzu",
    code == "AQ" ~ "Aqualog",
    code == "NU" ~ "CCAL",
    code == "LV" ~ "Levoglucosan",
    code == "EX" ~ "Excess",
    TRUE ~ NA_character_
  ))
datacodes

#sort the data by the randomn, since they should be filtered that way
datasortlab <- data_labels[order(data_labels$randomn),]
datasortlab #use this to print datasheets out


#repeat sample rows for each analysis
lab_data <- datasortlab %>% slice(rep(1:n(), each = analysisn)) %>%
  mutate(datacodes,
         lab_sample_ID =
           paste0(field_sample_ID, "_", datacodes$code ))
print(lab_data)
```

## Section 7- printing:

```{r warning=TRUE}

write_data.frames <- list(
  field_labels = data_labels %>% select(sample_name,field_sample_ID,botnum),
  lab_labels   = lab_data %>% select(sample_name,lab_sample_ID,analysis),
  lab_filter_sheets = data_labels[order(data_labels$randomn),] %>% select(sample_name,field_sample_ID,botnum),
  label_metadata = data_labels %>% select(site,datetime_PST,study,samp_type,randomn,botnum) #could delete this one  
)
safe.write_xlsx(write_data.frames,
                path = fileformat)
```
