---
title: "ISCO code"
author: "Emily Nussdorfer"
date: "2026-07-01"
---

Known time samples: Run code as usual

Unknown time samples: Run sections 1-4. After getting times, come back to section 5 and follow the instructions

## **Section 1- include sample information ([editable]{.underline}):**

Sample information 
```{r}
study <- "SSEMG" #the study code
samp_type <- "IS" #"IS" = ISCO
rn <- 1 #the start of a random number sequence. May change with projects or studies

```

```{r}
site <- c("STA","COA") #sites as a vector
nsamp <- c(24,24) #amount of samples per site. Matches the site vector above.
interval <- c(2,2) #interval of sampling ISCO is programmed for, usually 2hrs
blanktime <- "____________" # either have it blank  "____________" or with sampling date "YYYYMMDD____" or "YYYYMM______", etc
sc <- "20260223" #what sample campaign this is, the date it starts "YYYYMMDD" or something like sample_campaign-#. This is for file name purposes and doesn't effect sample naming
```


Analytes. Fill in TRUE or FALSE
```{r}
DOC <- TRUE
DOM <- TRUE
CCAL <- TRUE
Levoglucosan <- TRUE
Excess <- TRUE
```



```{r}
#optional change:
#can add a folder here if desired
#what format the files will end with. ex this returns: study_sapmpletype_date_label-table 
fileformat <- paste0(study,"_",samp_type,"_",sc,"_label-table.xlsx") #the file name
blank_fileformat <- paste0("blank_",study,"_",samp_type,"_",sc,"_label-table.csv") 
 
field_blank <- TRUE
lab_blank <- TRUE
nFB <- 1
nLB <- 1


```

```{r}
# Functions to safely write slsxs and csvs without overwriting anything
safe.write_xlsx <- function(x, path, ...) {
  if(file.exists(path)) {
    stop("The file you are trying to write already exists. DO NOT OVERWRITE UNLESS YOU MEAN TO!\n",
         "We don't want to overwrite randomized sample names\n",
         "Check ", path, " to see if the already written file is ok.")
  }
  write_xlsx(x, path, ...)
}


#' Title
#' 
#' Description here
#'
#' @param x 
#' @param path 
#' @param ... 
#'
#' @returns
#'
safe.write_csv <- function(x, path, ...) {
  if(file.exists(path)) {
    stop("The file you are trying to write already exists. DO NOT OVERWRITE UNLESS YOU MEAN TO!\n",
         "We don't want to overwrite randomized sample names\n",
         "Check ", path, " to see if the already written file is ok.")
  }
  write.csv(x, path, ...)
} #had to add this because couldn't get blank files to work wth xlsx




```


```{r}
sn <- sum(nsamp) #total sample amount from the vector that was filled out
#adding sample information
sampleinfo <- data.frame(site=site,nsamp=nsamp,interval=interval,study=study,samp_type=samp_type,blanktime=blanktime)
pt2 <- FALSE
```




## **Section 2- run libraries:**

```{r libraries, include=FALSE}
library(dplyr)
library(stringr)
library(lubridate)
library(writexl)
#source("helper-fun.R")
```

## **Section 3- code:**

Includes: total sample amount, and sample info, and functions to prevent overwrite files
```{r}
#FOR ISCO UNKNOWN

  #add sample names and IDs


data_label_final <- function(sampleinfo){
data_label <- function(sampleinfo){
    return(data.frame(site =rep(sampleinfo$site, times = sampleinfo$nsamp),
                      blank_time=sampleinfo$blanktime,
                      botnum = sprintf("%02d",rep(1:24, times= ceiling(sampleinfo$nsamp/24))[1:sampleinfo$nsamp]), study=sampleinfo$study,samp_type=sampleinfo$samp_type))
}
data_table <- lapply(1:nrow(sampleinfo),function(x){data_label(sampleinfo[x,])})%>%
bind_rows()


return(
  data.frame
  randomn=sample(rn:sn,sn, replace=F),
  sample_name = paste0( study, "_", "R", sprintf("%04d",randomn)),
  field_sample_ID_blank= paste0(study, "_", site, "_", samp_type,"_",blanktime))
)
}

# data_label <- function(sampleinfo){
#     return(data.frame(site =rep(sampleinfo$site, times = sampleinfo$nsamp),
#                       blank_time=sampleinfo$blanktime,
#                       botnum = sprintf("%02d",rep(1:24, times= ceiling(sampleinfo$nsamp/24))[1:sampleinfo$nsamp]), study=sampleinfo$study,samp_type=sampleinfo$samp_type))
#   }

randomn<-sample(rn:sn,sn, replace=F)
#where to add????
```

```{r pt 1, include=FALSE}

data_table <- lapply(1:nrow(sampleinfo),function(x){data_label(sampleinfo[x,])})%>%
bind_rows()%>%
mutate(
  randomn=sample(rn:sn,sn, replace=F),
  sample_name = paste0( study, "_", "R", sprintf("%04d",randomn)),
  field_sample_ID_blank= paste0(study, "_", site, "_", samp_type,"_",blanktime)
  )

data_names <- function(x){
  data<-x%>%
mutate(
  randomn=sample(rn:sn,sn, replace=F),
  sample_name = paste0( study, "_", "R", sprintf("%04d",randomn)),
  field_sample_ID_blank= paste0(study, "_", site, "_", samp_type,"_",blanktime)
  )

}

data_table <- lapply(1:nrow(sampleinfo),function(x){data_names(sampleinfo[x,])})%>%
bind_rows()#%>%

data_table
#this prints your data labels out if you have blank samples, since you will need to rerun this code again. 
#  safe.write_csv(data_table,
#                  path = blank_fileformat,
#                  row.names = FALSE)
# stop("DO NOT CONTINUE UNTIL YOU HAVE YOUR TIMES. SAVE YOUR CODE")



```

## **Section 5 - code** (for [blank times only]{.underline}):

*Note: can skip to section 6 if you have known times. It won't do anything for known time samples so running won't hurt*

Follow if you had blank times (this won't do anything for known times):

-   Do not change any of the above sections (leave known time FALSE)

-   Re-run sections 1-3 only

-   Fill the times and run code below ([editable]{.underline}):

```{r pt 5: and run, include=FALSE}
#enter the known start times on the ISCO (look at section 1 to match it to the correct site order)
start <- c("2026-02-23 15:00","2026-02-23 17:00")
pt2 <- TRUE
```

-   Run code below:

```{r pt 5: just run it, include=FALSE}

sampletimeinfo <- data.frame(nsamp=nsamp,interval=interval,start=start,
                             DOC=DOC,DOM=DOM,CCAL=CCAL,Levoglucosan=Levoglucosan, Excess=Excess)   

    site_label_blank <- function(sampletimeinfo){
      timestart <- as.POSIXct(sampletimeinfo$start)
      timelist <- seq(timestart, by=paste0(sampletimeinfo$interval, " hours"), length = sampletimeinfo$nsamp)
      # field_sample_ID <- paste0(study, "_", site, "_",samp_type,"_",datetime_PST)
      
      return(
        data.frame(time_PST=as.character(timelist),datetime_PST=format(timelist,"%Y%m%d%_%H%M"))
        #imports file using previous file naming#add sample Ids with time
       )

      }

# blank <- read.csv(blank_fileformat)
# blank
# %>%
#         mutate(field_sample_ID = paste0(study, "_",
#                                         site, "_",
#                                         samp_type,"_",
#                                         datetime_PST))%>%
#         cross_join(read.csv(blank_fileformat))

#make a data frame with times
data_labels <- lapply(1:nrow(sampletimeinfo),function(x){site_label_blank(sampletimeinfo[x,])}) %>% #makes the frame
  bind_rows()
data_labels
```

## **Section 6- adding lab analytes:**

```{r include=FALSE}



sampleinfo2 <- data.frame(site=site,nsamp=nsamp,start=start,interval=interval,study=study,samp_type=samp_type,blanktime=blanktime,DOC=DOC,DOM=DOM,CCAL=CCAL,Levoglucosan=Levoglucosan, Excess=Excess)

#sort the data by the randomn, since they should be filtered that way
datasortlab <- data_labels[order(data_labels$randomn),]
datasortlab #use this to print datasheets out




#adding in processing information:
#add analytes that are TRUE
analysis_opt <- c("CN", "AQ", "NU", "LV", "EX")
analysis_choice <- c(DOC, DOM, CCAL, Levoglucosan, Excess)

analysiscodes <- analysis_opt[analysis_choice]
analysiscodes
analysisn <- length(analysiscodes) #how many analytes there are
samples <- analysisn*sn #how many total filtered samples there will be

datacodes<- data.frame(code = rep(analysiscodes,times=sn)) %>% #add labcodes
  mutate(analysis = case_when(
    code == "CN" ~ "Shimadzu",
    code == "AQ" ~ "Aqualog",
    code == "NU" ~ "CCAL",
    code == "LV" ~ "Levoglucosan",
    code == "EX" ~ "Excess",
    TRUE ~ NA_character_
  ))
datacodes


#repeat sample rows for each analysis
lab_data <- datasortlab %>% slice(rep(1:n(), each = analysisn)) %>%
  mutate(datacodes,
         lab_sample_ID =
           paste0(field_sample_ID, "_", datacodes$code ))
print(lab_data)


labfunction <- function(
    
  
  
  
  
    )


```

## Section 7- printing:

```{r warning=TRUE}

write_data.frames <- list(
  # field_labels = data_labels %>% select(sample_name,field_sample_ID,botnum),
  lab_labels   = lab_data %>% select(sample_name,lab_sample_ID,analysis),
  lab_filter_sheets = data_labels[order(data_labels$randomn),] %>% select(sample_name,field_sample_ID),
  label_metadata = data_labels %>% select(site,datetime_PST,study,samp_type,randomn,botnum) #could delete this one  
)
safe.write_xlsx(write_data.frames,
                path = fileformat)
```
